Inferring disease architecture and predictive ability\\with LDpred2-auto

### 2022-10-05 paper/paper-infer.tex
typo

Most traits and diseases in humans are heritable.
What differs is the genetic architecture of each trait that can be parameterized by three key terms: the heritability (i.e.\ the proportion of phenotypic variation explained by genetics), the polygenicity (i.e.\ the fraction of genomic variants that have a non-zero impact on the trait), and the causal effect distribution (i.e.\ how the effect size distribution varies across causal variants).
Some phenotypes, such as schizophrenia or height, are highly heritable and highly polygenic \cite[]{sullivan2003schizophrenia,yang2010common,oconnor2019extreme}.
Causal effects are larger when a trait is more heritable, and smaller when it is more polygenic.
The distribution of causal effects relative to their allele frequencies is often investigated through a single parameter, usually called $\alpha$ or $S$, to model the effect of negative selection on complex traits whereby variants with lower frequencies are expected to have higher causal effect sizes \cite{speed2012improved}.
Many methods have been developed to estimate the SNP heritability (referred to as $h^2$ for brevity) and polygenicity ($p$), either globally for the whole genome or locally for specific regions of the genome, as well as $\alpha$. 
These methods include GCTA ($h^2$, \cite{yang2011gcta}), BOLT-REML ($h^2$ and $p$, \cite{loh2015contrasting}), LD Score regression ($h^2$, \cite{bulik2015ld}), FINEMAP (per-variant $p$ used for fine-mapping, \cite{benner2016finemap}), HESS (local $h^2$, \cite{shi2016contrasting}), LDAK-SumHer ($h^2$, \cite{speed2019sumher}), S-LD4M ($p$, \cite{oconnor2019extreme}), GRM-MAF-LD ($\alpha$, \cite{schoech2019quantification}), SuSiE (per-variant $p$ used for fine-mapping, \cite{wang2020simple}), SBayesS ($h^2$, $p$, and a third parameter $S$, similar to $\alpha$, \cite{zeng2021widespread}), and BEAVR (local $p$, \cite{johnson2021estimation}).

>>>>

Most traits and diseases in humans are heritable.
What differs is the genetic architecture of each trait that can be parameterized by three key terms: the heritability (i.e.\ the proportion of phenotypic variation explained by genetics), the polygenicity (i.e.\ the fraction of genomic variants that have a non-zero impact on the trait), and the causal effect distribution (i.e.\ how the effect size distribution varies across causal variants).
Some phenotypes, such as schizophrenia or height, are highly heritable and highly polygenic \cite[]{sullivan2003schizophrenia,yang2010common,oconnor2019extreme}.
Causal effects are larger when a trait is more heritable, and smaller when it is more polygenic.
The distribution of causal effects relative to their allele frequencies is often investigated through a single parameter, usually called $\alpha$ or $S$, to model the effect of negative selection on complex traits whereby variants with lower frequencies are expected to have higher causal effect sizes \cite[]{speed2012improved}.
Many methods have been developed to estimate the SNP heritability (referred to as $h^2$ for brevity) and polygenicity ($p$), either globally for the whole genome or locally for specific regions of the genome, as well as $\alpha$. 
These methods include GCTA ($h^2$, \cite{yang2011gcta}), BOLT-REML ($h^2$ and $p$, \cite{loh2015contrasting}), LD Score regression ($h^2$, \cite{bulik2015ld}), FINEMAP (per-variant $p$ used for fine-mapping, \cite{benner2016finemap}), HESS (local $h^2$, \cite{shi2016contrasting}), LDAK-SumHer ($h^2$, \cite{speed2019sumher}), S-LD4M ($p$, \cite{oconnor2019extreme}), GRM-MAF-LD ($\alpha$, \cite{schoech2019quantification}), SuSiE (per-variant $p$ used for fine-mapping, \cite{wang2020simple}), SBayesS ($h^2$, $p$, and a third parameter $S$, similar to $\alpha$, \cite{zeng2021widespread}), and BEAVR (local $p$, \cite{johnson2021estimation}).
